# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
1. Load the data
  ```{r}
  Sys.setlocale("LC_TIME", 'English')
  options(scipen=999)
  setwd('D:/Dropbox/Workshops&Courses/reproducible_research/RepData_PeerAssessment1')
  activity <- read.csv(unz('activity.zip', 'activity.csv'))
  ```

2. Transform the date column to the `Date` format in R. Also make the aggregated data set. 
  ```{r}
  activity <- transform(activity, date = as.Date(date, format = '%Y-%m-%d'))
  activity_agg1 <- aggregate(. ~ date, data = activity, FUN = sum)
  ```

## What is mean total number of steps taken per day?
1. Histogram:
  ```{r}
  hist(activity_agg1$steps)
  ```

  ```{r}
  mean1 <- mean(activity_agg1$steps)
  median1 <- median(activity_agg1$steps)
  ```

2. The mean total number of steps is `r mean1` and the median is `r median1`.

## What is the average daily activity pattern?
1. Time series plot (again using the `aggregate` function in R)
  ```{r}
  activity_agg2 <- aggregate(steps ~ interval, data = activity, FUN = mean)
  plot(steps ~ interval, data = activity_agg2, type = "l")
  ```

  ```{r}
  activity_agg2[which.max(activity_agg2$steps), ]
  ```

2. The interval between `r int <- activity_agg2[which.max(activity_agg2$steps), "interval"]; int` and `r int + 5` contains the maximum number of steps (`r activity_agg2[which.max(activity_agg2$steps), "steps"]`)

## Imputing missing values
  ```{r}
  sum(!complete.cases(activity))
  ```

1. The total number of missing values is `r sum(!complete.cases(activity))`. 

  ```{r}
  matching <- function(x) {
    count_NA <- sum(is.na(x))
    if (count_NA > 0) {
      imputed_values <- sample(x[!is.na(x)], count_NA, 
                              replace = TRUE)
      x[is.na(x)] <- imputed_values
    }
    return(x)
  }
  ```
  
2. I've written a simple function `matching` as above, where for each missing value, a number is randomly taken from other observations within the same 5-minute interval for imputation. 
  
  ```{r}
  activity_new <- within(activity, 
                         steps <- unlist(tapply(steps, 
                                                interval, 
                                                matching)))
  ```
3. The new dataset is `activity_new`.
  
4. Below shows the histogram

  ```{r}
  hist(with(activity_new, tapply(steps, date, sum)))
  mean(with(activity_new, tapply(steps, date, sum)))
  median(with(activity_new, tapply(steps, date, sum)))
  ```

  The mean is `r mean(with(activity_new, tapply(steps, date, sum)))` and the median is `r median(with(activity_new, tapply(steps, date, sum)))`.

The total number of missing values are `r sum(!complete.cases(activity))`. The values do not differ much from the estimates in the first part. 

## Are there differences in activity patterns between weekdays and weekends?
1. Below is the transformed data.

  ```{r}
  activity_new <- within(activity_new, weekday <- weekdays(date))
  activity_new <- within(activity_new, weekday <- !(weekday %in% c("Saturday", "Sunday")))
  activity_new <- within(activity_new, weekday <- factor(weekday, labels = c("weekend", "weekday")))
  summary(activity_new)
  ```
2. Below shows the plot.   
  
  ```{r}
  activity_agg1_new <- aggregate(. ~ interval + weekday, data = activity_new, FUN = mean)
  library(lattice)
  xyplot(steps ~ interval | weekday, data = activity_agg1_new, type = "l", 
  layout = c(1, 2), ylab = "Number of steps", xlab = "Interval")
  ```
